Meta-analysis

### 2020-06-17 03_Prep_data.Rmd
Changes to incorporate standardised mean differneces in the analysis

```{r}
slimmer_PM_dat %>%
   ggplot(aes(x = D_pres, y = grain_yield.t.ha))+
   geom_boxplot()

@@

>>>>

```{r plot_disease_pressure}
slimmer_PM_dat %>%
   ggplot(aes(x = D_pres, y = grain_yield.t.ha))+
   geom_boxplot()

@@

### Row spacing

We may want to consider rowspacing in the trials, However one of the locations has NA as the row spacing. 
Fow now lets assign the most used row spacing of 0.33
```{r}
slimmer_PM_dat[is.na(slimmer_PM_dat$row_spacing),]
slimmer_PM_dat[is.na(slimmer_PM_dat$row_spacing),"row_spacing"] <- 0.33

### 2020-06-17 04_MetaAnalysis.Rmd
Changes to incorporate standardised mean differneces in the analysis

The best model from the four above is `m11` with the lower AIC of -`r min(anova(m9, m11)$AIC)`

@@

### Imputing sample variances

We need to impute the variances which are missing for a few of the trials. These trials `r unique(slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),c("location","year", "trial_ref")])$trial_ref` were analysed and reported that there was no significant difference between the treatments in each trial. So therefore we need the imputed variances not to show a significant difference.

There are `r sum(is.na(slimmer_PM_dat$yield_error))` treatments without yield error in our data, and `r sum(!is.na(slimmer_PM_dat$yield_error))` where yield error was recorded.

Plotting a histogram of the variances show that the yield is not normally distributed. A log transformation, however, shows a normal distribution. We can use the mean and standard deviation of the log(V) to sample variances for the treatments where V is missing.

```{r variance_histograms}
hist_usq(slimmer_PM_dat$yield_error) # yield error is not normally distributed
hist_usq(log(slimmer_PM_dat$yield_error)) # a log transformation shows it is close to normally distributed 
```

We will evaluate whether the imputed variances confer a significant difference to the recorded means. I created a distance matrix of two standard errors above the mean against two standard errors below the mean. If any errors don't overlap, inferring significant difference, the imputed variances are discarded and re-sampled until all treatment variances show no significant differences.

```{r imputing_variances, eval=FALSE, include=FALSE}
source("./R/impute_yield_err.R")

imputed_error_list <- impute_yield_err(dat = slimmer_PM_dat, 
                                       xbar = "grain_yield.t.ha", 
                                       Var = "yield_error",
                                       reps = "n")

slimmer_PM_dat <- imputed_error_list[["new_dataFrame"]]

The imputed variances calculated above don't show the expected positive correlation between grain yield and variance. This is unexpected and reflects the inherent variability of the data. Therefore we will equalise the variance by using the trial mean squares and use this to calculate the variance for each sample.

@@

The use of these imputed MSE was giving a varied outcome each time it was run. We the next code chunk removes the studies without variation to observe what the meta-analysis outcome would be.

@@

@@

There seems like no difference between the treatments, with exception to Late_plus. Let's do this plot again, but use the proportion of yield increase compared to the no spray control as the response variable.

It is also important to visualise how well the data is compared across the trial years and trials.

@@

@@

```{r metafor_organisation}
slimmer_PM_dat$spray_management <-
   factor(
      slimmer_PM_dat$spray_management,
      c(
         "control",
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   )

slimmer_PM_dat$yi <- log(slimmer_PM_dat$grain_yield.t.ha)

slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
                                  2)

slimmer_PM_dat$trial_ref <- factor(slimmer_PM_dat$trial_ref)
slimmer_PM_dat$trial <- factor(slimmer_PM_dat$trial)
```

```{r Metafor-analysis}
PM_mv_AI <- rma.mv(
   yi,
   yield_error,
   mods = ~ spray_management,
   method = "ML",
   random = ~ 1|trial,
#   struct = "UN",
   data = slimmer_PM_dat
)

@@

According to the results of the omnibus test ($Q_M = 7.32, df = 5, p = 0.1979$) we can't reject the null hypothesis $H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$ that the tested spray management schedules have no benefit to protecting mungbean yields [@Viechtbauer2010]. 
However Late plus ($\beta_4$) was the only moderator showing a significant effect on yields ($z = 2.598; p = 0.0094$). 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 538.26, df = 141, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.
```{r metafor_contrasts}
meta_cont <- anova(PM_mv_AI, L = rbind(
   c(0, 1, -1, 0, 0,0),
   # early vs Recommended
   c(0, 1, 0, -1, 0,0),
   # early vs recommended plus
   c(0, 1, 0, 0, -1,0),
   # early vs Late 
   c(0, 1, 0, 0, 0,-1),
   # early vs Late plus
   c(0, 0, -1, 1, 0,0),
   # Recommended plus vs recommended
   c(0, 0, -1, 0, 1,0),
   # Late  vs Recommended
   c(0, 0, -1, 0, 0,1),
   # Late_plus  vs Recommended
   c(0, 0, 0, -1, 1,0),
   # Late vs Recommended plus
   c(0, 0, 0, -1, 0,1),
   # Late_plus  vs Recommended plus
   c(0, 0, 0, 0, -1,1)
)) # Late vs Late_plus

>>>>

There is no difference between the models, which means the the simpler model `m11` is the better model, also making it the best model from the four above with the lower AIC of -`r min(anova(m9, m11)$AIC)`

@@

@@

The use of these imputed MSE was giving a varied outcome each time it was run. 
The next code chunk removes the studies without variation to observe what the meta-analysis outcome would be.

@@

First, how well the data is compared across the trial years and trials.

@@

There seems like little difference between the treatments, with exception to Late_plus. 
Let's do this plot again, but lets look at the proportional mean difference between the treatments and the no spray control for each study.  
This should reduce variation in yield due between trials.

@@

Using the mean difference in the treatment effect seems to show the impact of each treatment better than just raw yield.
We know there is a good deal of variability in our studies and between them. 
Mungbean produces variable yields between seasons so we should use a response that highlights the difference in the treatment effects that we are interested in and reduce the variability.
Lets calculate the standardised mean difference for each treatment to reduce the variability in the meta-analysis.

### Caluculate standardised mean differences
We may wish to undertake the meta-analysis using standardised mean differences. 
Lets calculate them and add them to the dataframe.

#### Calculate sample variance from mean square errors

```{r MSE_2_SVar}
slimmer_PM_dat$vi <-
   slimmer_PM_dat$Y_Msquare / (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^
                                  2)
```

```{r}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for(T_ref in slimmer_PM_dat$trial_ref){

   # First we neee to Calculate the pooled standard deviation
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, "pooledSD"] <- 
   sqrt(sum((slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref,"n" ]-1) * slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref,"vi"])/
           (sum( slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref,"n"]) - nrow( slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref, ])))

   # Then Create a column with the value of the mean grain yield of the no spray control for comparisons to the treatments means
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref,"yi_C"] <-  
      mean(slimmer_PM_dat[slimmer_PM_dat$trial_ref == T_ref &
                        slimmer_PM_dat$fungicide_ai == "control",
                        "grain_yield.t.ha"], na.rm = TRUE)

}
# Calculate standardised mean difference
slimmer_PM_dat$grain_SMD <- (slimmer_PM_dat$grain_yield.t.ha - slimmer_PM_dat$yi_C)/ slimmer_PM_dat$pooledSD

@@

```{r metafor_organisation}
#remove controls from the data
dat1 <- slimmer_PM_dat[!slimmer_PM_dat$fungicide_ai == "control",]

dat1$spray_management <-
   factor(
      dat1$spray_management,
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   )

Lets inspect the data to determine if we need to transform the response variable

```{r}
hist(dat1$grain_yield.t.ha)
hist(log(dat1$grain_yield.t.ha))
hist(sqrt(dat1$grain_yield.t.ha))
hist(dat1$grain_SMD)
```
Standardised mean differences have a bit of a long tail however this is not bad and transformations with log or sqrt will not work on negative values.

Lets undertake the meta-analysis, we are using the `spray_management` variable as a moderator and a interactive term to the `trial` variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.
```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management -1,
   method = "ML",
   random = ~ spray_management|trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero. 
However the other treatments are significantly different to zero.
the Qm omnibus test of moderators show the moderators significantly influence the model and we can reject the null hypothesis that there is no difference between the moderators. 
The Qe test shoes there is a large significant amount of residual Heterogeneity.

Lets see if we can improve this model by including some other random effects and improving the random effect structure in the model.

#### CONTINUE FROM HERE!!!!!!!!!!

```{r}

PM_mv_AI2 <- rma.mv(
   yi,
   vi, # vi or yield error
   mods = ~ spray_management,
   method = "ML",
   random = ~ 1|trial,
   data = slimmer_PM_dat
)

anova(PM_mv_AI,PM_mv_AI2) # There is no significant difference between these models, Therefore use the simpler model with the lower AIC (PM_mv_AI2)

PM_mv_AI3 <- rma.mv(
   yi,
   vi, # vi or yield error
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1 | location/year/trial_ref, ~ 1 | host_genotype, ~ 1 | row_spacing),
   data = slimmer_PM_dat
)

anova(PM_mv_AI2,PM_mv_AI3) # There IS a significant difference between these models, Therefore use the more complicated model with the lower AIC (PM_mv_AI3)

PM_mv_AI4 <- rma.mv(
   yi,
   vi, # vi or yield error
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1 | location/year/trial_ref, ~ 1 | host_genotype),
   data = slimmer_PM_dat
)

anova(PM_mv_AI3,PM_mv_AI4) # There is no significant difference between these models, Therefore use the simpler model with the lower AIC (PM_mv_AI4)

profile(PM_mv_AI4, sigma2 = 1)

PM_mv_AI5 <- rma.mv(
   yi,
   vi, # vi or yield error
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1 | location/year/trial_ref),
   data = slimmer_PM_dat
)

anova(PM_mv_AI4,PM_mv_AI5) # There is no significant difference between these models, Therefore use the simpler model with the lower AIC (PM_mv_AI4)

summary(PM_mv_AI4)

PM_mv_AI6 <- rma.mv(
   yi,
   vi, # vi or yield error
   mods = ~ spray_management,
   method = "ML",
   random = list(~ 1 | location/year/trial_ref, ~ 1 | host_genotype, ~ spray_management | D_pres),
   data = slimmer_PM_dat
)

anova(PM_mv_AI6,PM_mv_AI4) # There is no significant difference between these models, Therefore use the simpler model with the lower AIC (PM_mv_AI4)

anova(PM_mv_AI2, PM_mv_AI3)

summary(PM_mv_AI2)

slimmer_PM_dat[is.na(slimmer_PM_dat$AUDPC_m),]
slimmer_PM_dat[slimmer_PM_dat$location == "Emerald", ]
table(slimmer_PM_dat$spray_management)
```
Lets try analysing the data as a risk reduction. 
For this we will use the proportion of yield reductions
```{r}
dat1 <- slimmer_PM_dat[!slimmer_PM_dat$fungicide_ai == "control",]

dat1$yi <- dat1$grain_SMD

PM_mv_MD <- rma.mv(
   yi = yi,
   vi, # vi or yield error
   mods = ~ spray_management - 1,
   method = "ML",
   random = list(~ 1 | location/year , ~1|host_genotype ,~ spray_management|D_pres),
   data = dat1
)

summary(PM_mv_MD)

dat1$D_pres

hist(dat1$grain_SMD)

@@

According to the results of the [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) ($Q_M = 7.32, df = 5, p = 0.1979$) we can't reject the null hypothesis $H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$ that the tested spray management schedules have no benefit to protecting mungbean yields [@Viechtbauer2010]. 
However Late plus ($\beta_4$) was the only moderator showing a significant effect on yields ($z = 2.598; p = 0.0094$). 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 538.26, df = 141, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

@@

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.
```{r metafor_contrasts}
meta_cont <- anova(PM_mv_MD, L = rbind(
   c(0, 1, -1, 0, 0,0),
   # early vs Recommended
   c(0, 1, 0, -1, 0,0),
   # early vs recommended plus
   c(0, 1, 0, 0, -1,0),
   # early vs Late 
   c(0, 1, 0, 0, 0,-1),
   # early vs Late plus
   c(0, 0, -1, 1, 0,0),
   # Recommended plus vs recommended
   c(0, 0, -1, 0, 1,0),
   # Late  vs Recommended
   c(0, 0, -1, 0, 0,1),
   # Late_plus  vs Recommended
   c(0, 0, 0, -1, 1,0),
   # Late vs Recommended plus
   c(0, 0, 0, -1, 0,1),
   # Late_plus  vs Recommended plus
   c(0, 0, 0, 0, -1,1)
)) # Late vs Late_plus
